---
title: "Week 10 Part 2"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    theme: sandstone
    number_sections: true
    code_folding: hide
---

```{r initial_setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(lubridate)
library(kableExtra)
library(scales)
knitr::opts_chunk$set(warning = FALSE)

theme_set(theme_bw())
```

# Chapter 15, Screening Many Models

What if we don't know which model to use?

## Workflow Sets

```{r}
data(concrete, package = "modeldata")
glimpse(concrete)

concrete <- 
   concrete %>% 
   group_by(across(-compressive_strength)) %>% 
   summarize(compressive_strength = mean(compressive_strength),
             .groups = "drop")
nrow(concrete)


set.seed(1501)
concrete_split <- initial_split(concrete, strata = compressive_strength)
concrete_train <- training(concrete_split)
concrete_test  <- testing(concrete_split)

set.seed(1502)
concrete_folds_mini <- 
   vfold_cv(concrete_train, strata = compressive_strength, repeats=1, v=3)

normalized_rec <- 
   recipe(compressive_strength ~ ., data = concrete_train) %>% 
   step_normalize(all_predictors()) 

poly_recipe <- 
   normalized_rec %>% 
   step_poly(all_predictors()) %>% 
   step_interact(~ all_predictors():all_predictors())
```


Some models (notably neural networks, K-nearest neighbors, and support vector machines) require predictors that have been centered and scaled, so some model workflows will require recipes with these preprocessing steps. For other models, a traditional response surface design model expansion (i.e., quadratic and two-way interactions) is a good idea. For these purposes, we create two recipes:

```{r}
library(rules)
library(baguette)

linear_reg_spec <- 
   linear_reg(penalty = tune(), mixture = tune()) %>% 
   set_engine("glmnet")

nnet_spec <- 
   mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>% 
   set_engine("nnet", MaxNWts = 2600) %>% 
   set_mode("regression")

mars_spec <- 
   mars(prod_degree = tune()) %>%  #<- use GCV to choose terms
   set_engine("earth") %>% 
   set_mode("regression")

svm_r_spec <- 
   svm_rbf(cost = tune(), rbf_sigma = tune()) %>% 
   set_engine("kernlab") %>% 
   set_mode("regression")

svm_p_spec <- 
   svm_poly(cost = tune(), degree = tune()) %>% 
   set_engine("kernlab") %>% 
   set_mode("regression")

knn_spec <- 
   nearest_neighbor(neighbors = tune(), dist_power = tune(), weight_func = tune()) %>% 
   set_engine("kknn") %>% 
   set_mode("regression")

cart_spec <- 
   decision_tree(cost_complexity = tune(), min_n = tune()) %>% 
   set_engine("rpart") %>% 
   set_mode("regression")

bag_cart_spec <- 
   bag_tree() %>% 
   set_engine("rpart", times = 50L) %>% 
   set_mode("regression")

rf_spec <- 
   rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
   set_engine("ranger") %>% 
   set_mode("regression")

xgb_spec <- 
   boost_tree(tree_depth = tune(), learn_rate = tune(), loss_reduction = tune(), 
              min_n = tune(), sample_size = tune(), trees = tune()) %>% 
   set_engine("xgboost") %>% 
   set_mode("regression")

cubist_spec <- 
   cubist_rules(committees = tune(), neighbors = tune()) %>% 
   set_engine("Cubist") 

nnet_param <- 
   nnet_spec %>% 
   extract_parameter_set_dials() %>% 
   update(hidden_units = hidden_units(c(1, 27)))
```


As a first workflow set example, let’s combine the recipe that only standardizes the predictors to the nonlinear models that require that the predictors be in the same units.

```{r}
normalized <- 
   workflow_set(
      preproc = list(normalized = normalized_rec), 
      models = list(SVM_radial = svm_r_spec, SVM_poly = svm_p_spec, 
                    KNN = knn_spec, neural_network = nnet_spec)
   )
normalized


normalized %>% extract_workflow(id = "normalized_KNN")


normalized <- 
   normalized %>% 
   option_add(param_info = nnet_param, id = "normalized_neural_network")
```


For the other nonlinear models, let’s create another workflow set that uses dplyr selectors for the outcome and predictors:


```{r}
model_vars <- 
   workflow_variables(outcomes = compressive_strength, 
                      predictors = everything())

no_pre_proc <- 
   workflow_set(
      preproc = list(simple = model_vars), 
      models = list(MARS = mars_spec, CART = cart_spec, CART_bagged = bag_cart_spec,
                    RF = rf_spec, boosting = xgb_spec, Cubist = cubist_spec)
   )
no_pre_proc


```


Finally, the set that uses nonlinear terms and interactions with the appropriate models are assembled:

```{r}
with_features <- 
   workflow_set(
      preproc = list(full_quad = poly_recipe), 
      models = list(linear_reg = linear_reg_spec, KNN = knn_spec)
   )
```


```{r}
all_workflows <- 
   bind_rows(no_pre_proc, normalized, with_features) %>% 
   # Make the workflow ID's a little more simple: 
   mutate(wflow_id = gsub("(simple_)|(normalized_)", "", wflow_id))
all_workflows
```

## Tuning


Almost all of these workflows contain tuning parameters. In order to evaluate their performance, we can use the standard tuning or resampling functions (e.g., `tune_grid()` and so on). The `workflow_map()` function will apply the same function to all of the workflows in the set; the default is `tune_grid()`. 

For this example, grid search is applied to each workflow using up to 25 different parameter candidates. There are a set of common options to use with each execution of `tune_grid()`. For example, we will use the same resampling and control objects for each workflow, along with a grid size of 25. The `workflow_map()` function has an additional argument called `seed` that is used to ensure that each execution of `tune_grid()` consumes the same random numbers. 

```{r}
grid_ctrl <-
   control_grid(
      save_pred = TRUE,
      parallel_over = "everything",
      save_workflow = TRUE
   )

grid_results <-
   all_workflows %>%
   workflow_map(
      seed = 1503,
      resamples = concrete_folds_mini,
      grid = 5,
      control = grid_ctrl,
      verbose = TRUE
   )

num_grid_models <- nrow(collect_metrics(grid_results, summarize = FALSE))
```


```{r}
grid_results %>% 
   rank_results() %>% 
   filter(.metric == "rmse") %>% 
   select(model, .config, rmse = mean, rank)
```

```{r}
autoplot(
   grid_results,
   rank_metric = "rmse",  # <- how to order models
   metric = "rmse",       # <- which metric to visualize
   select_best = TRUE     # <- one point per workflow
) +
   geom_text(aes(y = mean - 1/2, label = wflow_id), angle = 90, hjust = 1) +
   lims(y = c(3.5, 9.5)) +
   theme(legend.position = "none")
```

## Efficiently Screening Models

```{r}
library(finetune)

concrete_folds <- 
   vfold_cv(concrete_train, strata = compressive_strength, repeats=1, v=5)


race_ctrl <-
   control_race(
      save_pred = TRUE,
      parallel_over = "everything",
      save_workflow = TRUE
   )

race_results <-
   all_workflows %>%
   workflow_map(
      "tune_race_anova",
      seed = 1503,
      resamples = concrete_folds,
      grid = 5,
      control = race_ctrl,
      verbose = TRUE
   )
```

```{r}
autoplot(
   race_results,
   rank_metric = "rmse",  
   metric = "rmse",       
   select_best = TRUE    
) +
   geom_text(aes(y = mean - 1/2, label = wflow_id), angle = 90, hjust = 1) +
   lims(y = c(3.0, 9.5)) +
   theme(legend.position = "none")
```

## Finalize Model

```{r}
best_results <- 
   race_results %>% 
   extract_workflow_set_result("boosting") %>% 
   select_best(metric = "rmse")
best_results
```

```{r}
boosting_test_results <- 
   race_results %>% 
   extract_workflow("boosting") %>% 
   finalize_workflow(best_results) %>% 
   last_fit(split = concrete_split)
```

```{r}
collect_metrics(boosting_test_results)
```

```{r}
boosting_test_results %>% 
   collect_predictions() %>% 
   ggplot(aes(x = compressive_strength, y = .pred)) + 
   geom_abline(color = "gray50", lty = 2) + 
   geom_point(alpha = 0.5) + 
   coord_obs_pred() + 
   labs(x = "observed", y = "predicted")
```

